---
title: "Playing With Election Data"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 1
    #code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, cache=TRUE)
```
# Playing with Election Data - Looking at the starting boundaries + loading libraries

This is my first time playing around with RMarkdown, so first, I'm going to load in the data I'm working with. 

```{r results='hide', message=FALSE, warning=FALSE}

#loading in the libraries we need

library(leaflet)
library(sp)
library(rgdal)
library(maptools)
library(sf)

#Loading the ON Federal Election Districts 

E_D <- readOGR("ON_Electoral_Divisions.shp")
Poll_Boundaries <- readOGR("D:\\Blog_Personal_Work\\Hamilton_Election_Data\\Election_Poll_Boundaries\\Election_Poll_Boundaries.shp")

#Changing the projection of the District layer

E_Dist <- st_read("ON_Electoral_Divisions.shp") %>% 
  st_transform(crs="+init=epsg:4326")

#Just plotting the values normally

plot(E_D)

```

#Loading in data
Now that we have an idea of what the boundaries look like, lets load in a shapefile I merged in QGIS, indicating the 2015 federal election results for Ontario.

```{r}

#Selecting the appropriate colours for each political party 


pal <- colorFactor(
  palette = c('blue', 'red', 'orange'),
  domain = E_Dist$Federal__2
)

#Creating the map itself 
EDM <- leaflet(E_Dist) %>% 
  addProviderTiles(providers$OpenStreetMap) %>%
  setView(-80,49,4) %>%
  addPolygons(data=E_Dist,
              weight = 2,
              opacity = 1,
              color = ~pal(Federal__2),
              dashArray = "3",
              fillOpacity = 0.7,
              highlight = highlightOptions(
                weight = 5,
                color = "#666",
                dashArray = "",
                fillOpacity = 0.7,
                bringToFront = TRUE),
                label = ~Federal_12)
                

#And done! Lets take a look. 

EDM
```


# Looking specifically at the Greater Toronto Area

Awesome, once we've seen the entirety of the province, let's specifically look at the Greater Toronto Area

```{R }
EDM_GTA <- leaflet(E_Dist) %>% 
  addProviderTiles(providers$OpenStreetMap) %>%
  setView(-79.5,43.6,9) %>%
  addPolygons(data=E_Dist,
              weight = 2,
              opacity = 1,
              color = ~pal(Federal__2),
              dashArray = "3",
              fillOpacity = 0.7,
              highlight = highlightOptions(
                weight = 5,
                color = "#666",
                dashArray = "",
                fillOpacity = 0.7,
                bringToFront = TRUE),
                label = ~Federal_12)
                

# Now lets get a closer look of the GTA 
EDM_GTA


```

And done! 


